---
title: "Latex for hypothesis testing"
author: "Jeffrey Liang"
date: "10/29/2020"
output: 
  html_document:
    toc: true
    toc_float : true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = F)
```

R Code can be found in [here](./rcode.html)

# Distribution {.tabset}

## Binomial
$$
P(\mathcal{X}=k)\ = {n \choose k}p^k(1-p)^{n-k}
$$
```{r}
$P(\mathcal{X}=k)\ = {n \choose k}p^k(1-p)^{n-k}$
```


## Normal Distribution

$P(\mathcal{X}=k)\ = \frac{1}{\sqrt{2\pi}\sigma}*e^{-\frac{(x-\mu)^2}{2\sigma^2}}$
```{r}
$P(\mathcal{X}=k)\ = \frac{1}{\sqrt{2\pi}\sigma}*e^{-\frac{(x-\mu)^2}{2\sigma^2}}$
```



# One Group t-test {.tabset}

## left tailed

$H_0$: there's no difference betwee sample mean and the true mean

$H_1$: the sample mean is lower than the true mean

$\bar{x} = \frac{\sum_{i=1}^n{x_i}}{n}$

$s_d = \sqrt{\sum_{i=1}^n{(x_i - \bar{x})^2}/(n-1)}$

$t = \frac{\bar{x} - 0}{s_d/\sqrt{n}}$

$Reject ~ H_0 ~ if ~ t<t_{n-1,\alpha}$

$Fail ~ reject ~ H_0 ~ if ~ t>t_{n-1,\alpha}$

$CI : (-\infty,\bar{x} - t_{df,\alpha}*sd/\sqrt{n})$

```{r}
$H_0$: there's no difference betwee sample mean and the true mean

$H_1$: the sample mean is lower than the true mean

$\bar{x} = \frac{\sum_{i=1}^n{x_i}}{n}$

$s_d = \sqrt{\sum_{i=1}^n{(x_i - \bar{x})^2}/(n-1)}$

$t = \frac{\bar{x} - 0}{s_d/\sqrt{n}}$

$Reject ~ H_0 ~ if ~ t<t_{n-1,\alpha}$

$Fail ~ reject ~ H_0 ~ if ~ t>t_{n-1,\alpha}$

$CI : (-\infty,\bar{x} - t_{df,\alpha}*sd/\sqrt{n})$
```


## right tailed

$H_0$: there's no difference betwee sample mean and the true mean

$H_1$: the sample mean is greater than the true mean

$\bar{x} = \frac{\sum_{i=1}^n{x_i}}{n}$

$s_d = \sqrt{\sum_{i=1}^n{(x_i - \bar{x})^2}/(n-1)}$

$t = \frac{\bar{x} - 0}{s_d/\sqrt{n}}$

$Reject ~ H_0 ~ if ~ t>t_{n-1,1-\alpha}$

$Fail ~ reject ~ H_0 ~ if ~ t<t_{n-1,1-\alpha}$

$CI : (\bar{x} + t_{df,1-\alpha}*sd/\sqrt{n},\infty)$

```{r}
$H_0$: there's no difference betwee sample mean and the true mean

$H_1$: the sample mean is greater than the true mean

$\bar{x} = \frac{\sum_{i=1}^n{x_i}}{n}$

$s_d = \sqrt{\sum_{i=1}^n{(x_i - \bar{x})^2}/(n-1)}$

$t = \frac{\bar{x} - 0}{s_d/\sqrt{n}}$

$Reject ~ H_0 ~ if ~ t>t_{n-1,1-\alpha}$

$Fail ~ reject ~ H_0 ~ if ~ t<t_{n-1,1-\alpha}$

$CI : (\bar{x} + t_{df,1-\alpha}*sd/\sqrt{n},\infty)$
```


## two-side

$H_0$: there's no difference betwee sample mean and the true mean

$H_1$: the sample mean is different from the true mean

$\bar{x} = \frac{\sum_{i=1}^n{x_i}}{n}$

$s_d = \sqrt{\sum_{i=1}^n{(x_i - \bar{x})^2}/(n-1)}$

$t = \frac{\bar{x} - 0}{s_d/\sqrt{n}}$

$Reject ~ H_0 ~ if ~ |t|<t_{n-1,1-\alpha/2}$

$Fail ~ reject ~ H_0 ~ if ~ |t|<t_{n-1,1 - \alpha/2}$

$CI : (\bar{x} - t_{df,1-\alpha/2}*sd/\sqrt{n} ,\bar{x} + t_{df,1-\alpha/2}*sd/\sqrt{n})$


```{r}
$H_0$: there's no difference betwee sample mean and the true mean

$H_1$: the sample mean is different from the true mean

$\bar{x} = \frac{\sum_{i=1}^n{x_i}}{n}$

$s_d = \sqrt{\sum_{i=1}^n{(x_i - \bar{x})^2}/(n-1)}$

$t = \frac{\bar{x} - 0}{s_d/\sqrt{n}}$

$Reject ~ H_0 ~ if ~ |t|<t_{n-1,1-\alpha/2}$

$Fail ~ reject ~ H_0 ~ if ~ |t|<t_{n-1,1 - \alpha/2}$

$CI : (\bar{x} - t_{df,1-\alpha/2}*sd/\sqrt{n} ,\bar{x} + t_{df,1-\alpha/2}*sd/\sqrt{n})$

```

### paired
$H_0$: there's no difference betwee difference

$H_1$: the difference is different

$\bar{d} = \frac{\sum_{i=1}^n{d_i}}{n}$

$s_d = \sqrt{\sum_{i=1}^n{(d_i - \bar{d})^2}/(n-1)}$

$t = \frac{\bar{d} - 0}{s_d/\sqrt{n}}$

$Reject ~ H_0 ~ if ~ |t|<t_{n-1,1-\alpha/2}$

$Fail ~ reject ~ H_0 ~ if ~ |t|<t_{n-1,1 - \alpha/2}$

$CI : (\bar{d} - t_{df,1-\alpha/2}*sd/\sqrt{n} ,\bar{d} + t_{df,1-\alpha/2}*sd/\sqrt{n})$


```{r}
$H_0$: there's no difference betwee difference

$H_1$: the difference is different

$\bar{d} = \frac{\sum_{i=1}^n{d_i}}{n}$

$s_d = \sqrt{\sum_{i=1}^n{(d_i - \bar{d})^2}/(n-1)}$

$t = \frac{\bar{d} - 0}{s_d/\sqrt{n}}$

$Reject ~ H_0 ~ if ~ |t|<t_{n-1,1-\alpha/2}$

$Fail ~ reject ~ H_0 ~ if ~ |t|<t_{n-1,1 - \alpha/2}$

$CI : (\bar{x} - t_{df,1-\alpha/2}*sd/\sqrt{n} ,\bar{d} + t_{df,1-\alpha/2}*sd/\sqrt{n})$
```


# Two Group testing {.tabset}

## variance test

$H_0$: the variances between group are equal(no difference)
    
$H_1$: the variances between group are not equal(there's difference)

$s_{x_1} = \sqrt{\sum_{i=1}^{n_1}(x_i - \bar{x_1})^2/(n_1-1)}$

$s_{x_2} = \sqrt{\sum_{j=1}^{n_2}(x_i - \bar{x_2})^2/(n_2-1)}$

$F = s_1^2/s_2^2  \sim F_{n_1-1,n_2-1}$

$Reject ~ H_0 ~ if ~ F>F_{n_1-1,n_2-1,1-\alpha/2} ~ OR ~ F<F_{n_1-1,n_2-1,\alpha/2}$

$Fail ~ reject ~ H_0 ~ if ~ F_{n_1-1,n_2-1,\alpha/2}<F<F_{n_1-1,n_2-1,1-\alpha/2}$

```{r}
$H_0$: the variances between group are equal(no difference)
    
$H_1$: the variances between group are not equal(there's difference)

$s_{x_1} = \sqrt{\sum_{i=1}^{n_1}(x_i - \bar{x_1})^2/(n_1-1)}$

$s_{x_2} = \sqrt{\sum_{j=1}^{n_2}(x_i - \bar{x_2})^2/(n_2-1)}$

$F = s_1^2/s_2^2  \sim F_{n_1-1,n_2-1}$

$Reject ~ H_0 ~ if ~ F>F_{n_1-1,n_2-1,1-\alpha/2} ~ OR ~ F<F_{n_1-1,n_2-1,\alpha/2}$

$Fail ~ reject ~ H_0 ~ if ~ F_{n_1-1,n_2-1,\alpha/2}<F<F_{n_1-1,n_2-1,1-\alpha/2}$
```

## t-test with equal variance
$H_0$: the means between group are equal(no difference)
    
$H_1$: the means between group are not equal(there's difference)

$s_{pool} = \frac{(n_1-1)s_1 + (n_2 -1)s_2}{n_1+n_2-2}$

$t = \frac{\bar{X_1} - \bar{X_2}}{s_{pool}\times\sqrt{\frac{1}{n_1}+\frac{1}{n_2}}}$

$Reject ~ H_0 ~ if ~ |t|>t_{df,1-\alpha/2}$

$Fail ~ reject ~ H_0 ~ if ~ |t|<t_{df,1-\alpha/2}$

$CI = (\bar{X_1}-\bar{X_2} ~ - ~ t_{df,1-\alpha/2}s_{pool}/\sqrt{1/n_1+1/n_2},\bar{X_1}-\bar{X_2} ~ + ~ t_{df,1-\alpha/2}s_{pool}/\sqrt{1/n_1+1/n_2} )$

```{r}
$H_0$: the means between group are equal(no difference)
    
$H_1$: the means between group are not equal(there's difference)

$s_{pool} = \frac{(n_1-1)s_1 + (n_2 -1)s_2}{n_1+n_2-2}$

$t = \frac{\bar{X_1} - \bar{X_2}}{s_{pool}\times\sqrt{\frac{1}{n_1}+\frac{1}{n_2}}}$

$Reject ~ H_0 ~ if ~ |t|>t_{df,1-\alpha/2}$

$Fail ~ reject ~ H_0 ~ if ~ |t|<t_{df,1-\alpha/2}$

$CI = (\bar{X_1}-\bar{X_2} ~ - ~ t_{df,1-\alpha/2}s_{pool}/\sqrt{1/n_1+1/n_2},\bar{X_1}-\bar{X_2} ~ + ~ t_{df,1-\alpha/2}s_{pool}/\sqrt{1/n_1+1/n_2} )$
```

## t-test with unequal variance
$H_0$: the means between group are equal(no difference)
    
$H_1$: the means between group are not equal(there's difference)

$\bar{X_1} - \bar{X_2} \sim N(\mu_1-\mu_2,\frac{\sigma_1^2}{n_1}+\frac{\sigma_2^2}{n_2})$ if we know the population variance

$t = \frac{\bar{X_1} - \bar{X_2}}{\sqrt{\frac{s_1^2}{n_1}+\frac{s_2^2}{n_2}}}~\sim t_{d''}$

$d' = round(d'') = \frac{(\frac{s_1^2}{n_1}+\frac{s_2^2}{n_2})^2}{\frac{s_1^2}{n_1}^2/(n_1-1)+\frac{s_2^2}{n_2}^2/(n_2-1)}$

$Reject ~ H_0 ~ if ~ |t|>t_{df,1-\alpha/2}$

$Fail ~ reject ~ H_0 ~ if ~ |t|<t_{df,1-\alpha/2}$

$CI = (\bar{X_1}-\bar{X_2} ~ - ~ t_{df,1-\alpha/2}\sqrt{\frac{s_1^2}{n_1}+\frac{s_2^2}{n_2}},\bar{X_1}-\bar{X_2} ~ + ~ t_{df,1-\alpha/2}\sqrt{\frac{s_1^2}{n_1}+\frac{s_2^2}{n_2}})$

```{r}
$H_0$: the means between group are equal(no difference)
    
$H_1$: the means between group are not equal(there's difference)

$\bar{X_1} - \bar{X_2} \sim N(\mu_1-\mu_2,\frac{\sigma_1^2}{n_1}+\frac{\sigma_2^2}{n_2})$ if we know the population variance

$t = \frac{\bar{X_1} - \bar{X_2}}{\sqrt{\frac{s_1^2}{n_1}+\frac{s_2^2}{n_2}}}~\sim t_{d''}$

$d' = round(d'') = \frac{(\frac{s_1^2}{n_1}+\frac{s_2^2}{n_2})^2}{\frac{s_1^2}{n_1}^2/(n_1-1)+\frac{s_2^2}{n_2}^2/(n_2-1)}$

$Reject ~ H_0 ~ if ~ |t|>t_{df,1-\alpha/2}$

$Fail ~ reject ~ H_0 ~ if ~ |t|<t_{df,1-\alpha/2}$

$CI = (\bar{X_1}-\bar{X_2} ~ - ~ t_{df,1-\alpha/2}\sqrt{\frac{s_1^2}{n_1}+\frac{s_2^2}{n_2}},\bar{X_1}-\bar{X_2} ~ + ~ t_{df,1-\alpha/2}\sqrt{\frac{s_1^2}{n_1}+\frac{s_2^2}{n_2}})$
```

# Multigroup Camparison {.tabset}

## ANOVA
$H_0$ : there's no difference between groups

$H_1$ : at least one group is different from the other groups

$Between~Sum~of~Square = \sum_{i=1}^k\sum_{j=1}^{n_i}(\bar{y_i} - \bar{\bar{y}})^2=\sum_i^kn_i\bar{y_i}^2-\frac{y_{..}^2}{n}$

$Within~Sum~of~Square = \sum_{i=1}^k\sum_{j=1}^{n_i}(y_{ij}-\bar{y_i})^2=\sum_i^k(n_i-1)s_i^2$

$Between~Mean~Square = \frac{\sum_{i=1}^k\sum_{j=1}^{n_i}(\bar{y_i} - \bar{\bar{y}})^2}{k-1}$

$Within~Mean~Square = \frac{\sum_{i=1}^k\sum_{j=1}^{n_i}(y_{ij}-\bar{y_i})^2}{n-k}$

$F_{statistics} = \frac{Between~Mean~Square}{Within~Mean~Square} \sim F(k-1,n-k)$

$Reject ~ H_0 ~ if ~ F>F_{k-1,n-k,1-\alpha}$

$Fail ~ reject ~ H_0 ~ if ~F<F_{k-1,n-k,1-\alpha}$
```{r}
$H_0$ : there's no difference between groups

$H_1$ : at least one group is different from the other groups

$Between~Sum~of~Square = \sum_{i=1}^k\sum_{j=1}^{n_i}(\bar{y_i} - \bar{\bar{y}})^2=\sum_i^kn_i\bar{y_i}^2-\frac{y_{..}^2}{n}$

$Within~Sum~of~Square = \sum_{i=1}^k\sum_{j=1}^{n_i}(y_{ij}-\bar{y_i})^2=\sum_i^k(n_i-1)s_i^2$

$Between~Mean~Square = \frac{\sum_{i=1}^k\sum_{j=1}^{n_i}(\bar{y_i} - \bar{\bar{y}})^2}{k-1}$

$Within~Mean~Square = \frac{\sum_{i=1}^k\sum_{j=1}^{n_i}(y_{ij}-\bar{y_i})^2}{n-k}$

$F_{statistics} = \frac{Between~Mean~Square}{Within~Mean~Square} \sim F(k-1,n-k)$

$Reject ~ H_0 ~ if ~ F>F_{k-1,n-k,1-\alpha}$

$Fail ~ reject ~ H_0 ~ if ~F<F_{k-1,n-k,1-\alpha}$
```

# Proportion testing {.tabset}

## Normal Approximation

## Homogeneity Chi-sq test

$H_0 :p_{1j} =p_{2j}=...=p_{ij}$ the proportion among $group_i$ are equal ...

$H_1$ : For at least one column there're two row i and i' where the proability are not the same.

$\mathcal{X}^2 = \sum_i^{row}\sum_j^{col}\frac{(n_{ij}-E_{ij})^2}{E_{ij}} \sim \mathcal{X}^2_{df = (row-1)\times(col_1)}$

$Reject ~ H_0 ~ if ~ \mathcal{X}^2>\mathcal{X}^2_{(r-1))*(c-1),1-\alpha}$

$Fail ~ reject ~ H_0 ~ if ~\mathcal{X}^2<\mathcal{X}^2_{(r-1))*(c-1),1-\alpha}$

```{r}
$H_0$ :p_{1j} =p_{2j}=...=p_{ij}$ the proportion among $group_i$ are equal ...

$H_1$ : For at least one column there're two row i and i' where the proability are not the same.

$\mathcal{X}^2 = \sum_i^{row}\sum_j^{col}\frac{(n_{ij}-E_{ij})^2}{E_{ij}} \sim \mathcal{X}^2_{df = (row-1)\times(col-1)}$
  
$Reject ~ H_0 ~ if ~ \mathcal{X}^2>\mathcal{X}^2_{(r-1))*(c-1),1-\alpha}$

$Fail ~ reject ~ H_0 ~ if ~\mathcal{X}^2<\mathcal{X}^2_{(r-1))*(c-1),1-\alpha}$
```


## Independent test

$H_0$ : Group A and Group B are independet

$H_1$ : Group A and Group B are dependet/associate

$\mathcal{X}^2 = \sum_i^{row}\sum_j^{col}\frac{(n_{ij}-E_{ij})^2}{E_{ij}} \sim \mathcal{X}^2_{df = (row-1)\times(col-1)}$
  
$Reject ~ H_0 ~ if ~ \mathcal{X}^2>\mathcal{X}^2_{(r-1))*(c-1),1-\alpha}$

$Fail ~ reject ~ H_0 ~ if ~\mathcal{X}^2<\mathcal{X}^2_{(r-1))*(c-1),1-\alpha}$

```{r}
$H_0$ : Group A and Group B are independet

$H_1$ : Group A and Group B are dependet/associate

$\mathcal{X}^2 = \sum_i^{row}\sum_j^{col}\frac{(n_{ij}-E_{ij})^2}{E_{ij}} \sim \mathcal{X}^2_{df = (row-1)\times(col-1)}$
  
$Reject ~ H_0 ~ if ~ \mathcal{X}^2>\mathcal{X}^2_{(r-1))*(c-1),1-\alpha}$

$Fail ~ reject ~ H_0 ~ if ~\mathcal{X}^2<\mathcal{X}^2_{(r-1))*(c-1),1-\alpha}$
  
```

## Fisher Exact

## McNemar Test

