---
title: "This is the site for latex for hypothesis testing"
author: "Jeffrey Liang"
date: "10/29/2020"
output: 
  html_document:
    toc: true
    toc_float : true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Distribution

### Binomial
$$
P(\mathcal{X}=k)\ = {n \choose k}p^k(1-p)^{n-k}
$$
```{r}
#$P(\mathcal{X}=k)\ = {n \choose k}p^k(1-p)^{n-k}$
```


### Normal Distribution

$P(\mathcal{X}=k)\ = \frac{1}{\sqrt{2\pi}\sigma}*e^{-\frac{(x-\mu)^2}{2\sigma^2}}$
```{r}
#$P(\mathcal{X}=k)\ = \frac{1}{\sqrt{2\pi}\sigma}*e^{-\frac{(x-\mu)^2}{2\sigma^2}}$
```



# One Group Testing

## t-test 

### left tailed

$H_0$ : there's no difference betwee sample mean and the true mean

$H_1$ : the sample mean is lower than the true mean

$\bar{x} = \frac{\sum_{i=1}^n{x_i}}{n}$

$s_d = \sqrt{\sum_{i=1}^n{(x_i - \bar{x})^2}/(n-1)}$

$t = \frac{\bar{x} - 0}{s_d/\sqrt{n}}$

$Reject~H_0~if~t<t_{n-1,\alpha}$

$Fail~reject~H_0~if~t>t_{n-1,\alpha}$

$CI : (-\infty,\bar{x} - t_{df,\alpha}*sd/\sqrt{n})$ 

```{r}
#$H_0$ : there's no difference betwee sample mean and the true mean

#$H_1$ : the sample mean is lower than the true mean

#$\bar{x} = \frac{\sum_{i=1}^n{x_i}}{n}$

#$s_d = \sqrt{\sum_{i=1}^n{(x_i - \bar{x})^2}/(n-1)}$

#$t = \frac{\bar{x} - 0}{s_d/\sqrt{n}}$

#$Reject~H_0~if~t<t_{n-1,\alpha}$

#$Fail~reject~H_0~if~t>t_{n-1,\alpha}$

#$CI : (-\infty,\bar{x} - t_{df,\alpha}*sd/\sqrt{n})$ 
```


### right tailed

```{r}
#$H_0$ : there's no difference betwee sample mean and the true mean

#$H_1$ : the sample mean is greater than the true mean

#$\bar{x} = \frac{\sum_{i=1}^n{x_i}}{n}$

#$s_d = \sqrt{\sum_{i=1}^n{(x_i - \bar{x})^2}/(n-1)}$

#$t = \frac{\bar{x} - 0}{s_d/\sqrt{n}}$

#$Reject~H_0~if~t>t_{n-1,1-\alpha}$

#$Fail~reject~H_0~if~t<t_{n-1,1-\alpha}$

#$CI : (\bar{x} + t_{df,1-\alpha}*sd/\sqrt{n},\infty)$
```


### two-side

```{r}
#$H_0$ : there's no difference betwee sample mean and the true mean

#$H_1$ : the sample mean is different from the true mean

#$\bar{x} = \frac{\sum_{i=1}^n{x_i}}{n}$

#$s_d = \sqrt{\sum_{i=1}^n{(x_i - \bar{x})^2}/(n-1)}$

#$t = \frac{\bar{x} - 0}{s_d/\sqrt{n}}$

#$Reject~H_0~if~|t|<t_{n-1,1-\alpha/2}$

#$Fail~reject~H_0~if~|t|<t_{n-1,1 - \alpha/2}$

#$CI : (\bar{x} - t_{df,1-\alpha/2}*sd/\sqrt{n} ,\bar{x} - t_{df,1-\alpha/2}*sd/\sqrt{n})$

```


# Two Group testing

## t-test

### paired
$H_0$ : there's no difference betwee difference

$H_1$ : the difference is different

$\bar{d} = \frac{\sum_{i=1}^n{d_i}}{n}$

$s_d = \sqrt{\sum_{i=1}^n{(d_i - \bar{d})^2}/(n-1)}$

$t = \frac{\bar{d} - 0}{s_d/\sqrt{n}}$

$Reject~H_0~if~|t|<t_{n-1,1-\alpha/2}$

$Fail~reject~H_0~if~|t|<t_{n-1,1 - \alpha/2}$

$CI : (\bar{d} - t_{df,1-\alpha/2}*sd/\sqrt{n} ,\bar{d} - t_{df,1-\alpha/2}*sd/\sqrt{n})$


```{r}
#$H_0$ : there's no difference betwee difference

#$H_1$ : the difference is different

#$\bar{d} = \frac{\sum_{i=1}^n{d_i}}{n}$

#$s_d = \sqrt{\sum_{i=1}^n{(d_i - \bar{d})^2}/(n-1)}$

#$t = \frac{\bar{d} - 0}{s_d/\sqrt{n}}$

#$Reject~H_0~if~|t|<t_{n-1,1-\alpha/2}$

#$Fail~reject~H_0~if~|t|<t_{n-1,1 - \alpha/2}$

#$CI : (\bar{x} - t_{df,1-\alpha/2}*sd/\sqrt{n} ,\bar{d} - t_{df,1-\alpha/2}*sd/\sqrt{n})$
```

### paried test
$H_0$ : there's no difference betwee two groups

$H_1$ : the means of groups' are different

$\bar{x_1} = \frac{\sum_{i=1}^n{x_{1i}}}{n}$

$\bar{x_2} = \frac{\sum_{i=1}^n{x_{2i}}}{n}$

$s_d = \sqrt{\sum_{i=1}^n{(d_i - \bar{d})^2}/(n-1)}

$t = \frac{\bar{d} - 0}{s_d/\sqrt{n}}

$Reject~H_0~if~|t|<t_{n-1,1-\alpha/2}

$Fail~reject~H_0~if~|t|<t_{n-1,1 - \alpha/2}$

$CI : (\bar{d} - t_{df,1-\alpha/2}*sd/\sqrt{n} ,\bar{d} - t_{df,1-\alpha/2}*sd/\sqrt{n})$

```{r}
#$H_0$ : there's no difference betwee two groups

#$H_1$ : the means of groups' are different

#$\bar{x_1} = \frac{\sum_{i=1}^n{x_{1i}}}{n}$

#$\bar{x_2} = \frac{\sum_{i=1}^n{x_{2i}}}{n}$

#$s_d = \sqrt{\sum_{i=1}^n{(d_i - \bar{d})^2}/(n-1)}

#$t = \frac{\bar{d} - 0}{s_d/\sqrt{n}}

#$Reject~H_0~if~|t|<t_{n-1,1-\alpha/2}

#$Fail~reject~H_0~if~|t|<t_{n-1,1 - \alpha/2}$

#$CI : (\bar{d} - t_{df,1-\alpha/2}*sd/\sqrt{n} ,\bar{d} - t_{df,1-\alpha/2}*sd/\sqrt{n})$
```


## Two group

### variance test

$H_0$: the variances between group are the equal(no difference)
    
$H_1$: the variances between group are not the equal(there's difference)

$s_{x_1} = \sqrt{\sum_{i=1}^{n_1}(x_i - \bar{x_1})^2/(n_1-1)}$

$s_{x_2} = \sqrt{\sum_{j=1}^{n_2}(x_i - \bar{x_2})^2/(n_2-1)}$

$F = s_1^2/s_2^2  \sim F_{n_1-1,n_2-1}$

$ Reject~H_0~if~F>F_{n_1-1,n_2-1,1-\alpha/2}~OR~F<F_{n_1-1,n_2-1,\alpha/2}$

$ Fail~reject~H_0~if~F_{n_1-1,n_2-1,\alpha/2}<F<F_{n_1-1,n_2-1,1-\alpha/2}$

```{r}
#$H_0$: the variances between group are the equal(no difference)
    
#$H_1$: the variances between group are not the equal(there's difference)

#$s_{x_1} = \sqrt{\sum_{i=1}^{n_1}(x_i - \bar{x_1})^2/(n_1-1)}$

#$s_{x_2} = \sqrt{\sum_{j=1}^{n_2}(x_i - \bar{x_2})^2/(n_2-1)}$

#$F = s_1^2/s_2^2  \sim F_{n_1-1,n_2-1}$

#$ Reject~H_0~if~F>F_{n_1-1,n_2-1,1-\alpha/2}~OR~F<F_{n_1-1,n_2-1,\alpha/2}$

#$ Fail~reject~H_0~if~F_{n_1-1,n_2-1,\alpha/2}<F<F_{n_1-1,n_2-1,1-\alpha/2}$
```

### t-test with equal variance

$s_{pool} = \frac{(n_1-1)s_1 + (n_2 -1)s_2}{n_1+n_2-2}$

$t = \frac{\bar{X_1} - \bar{x_2}}{s_{pool}\times\sqrt{\frac{1}{n_1}+\frac{1}{n_2}}}$

$ Reject~H_0~if~|t|>t_{df,1-\alpha/2}$

$ Fail~reject~H_0~if~|t|<t_{df,1-\alpha/2}$

$CI = (\bar{X_1}-\bar{x_2}~-~t_{df,1-\alpha/2}s_{pool}/\sqrt{1/n_1+1/n_2},\bar{X_1}-\bar{x_2}~-~t_{df,1-\alpha/2}s_{pool}/\sqrt{1/n_1+1/n_2} )$

```{r}
#$s_{pool} = \frac{(n_1-1)s_1 + (n_2 -1)s_2}{n_1+n_2-2}$

#$t = \frac{\bar{X_1} - \bar{x_2}}{s_{pool}\times\sqrt{\frac{1}{n_1}+\frac{1}{n_2}}}$

#$ Reject~H_0~if~|t|>t_{df,1-\alpha/2}$

#$ Fail~reject~H_0~if~|t|<t_{df,1-\alpha/2}$

#$CI = (\bar{X_1}-\bar{x_2}~-~t_{df,1-\alpha/2}s_{pool}/\sqrt{1/n_1+1/n_2},\bar{X_1}-\bar{x_2}~-~t_{df,1-\alpha/2}s_{pool}/\sqrt{1/n_1+1/n_2} )$
```

