---
output:
  html_document:
    df_print: paged
header-includes: \usepackage{graphicx} \usepackage{float} \usepackage{amsmath}
---

```{r setup, include = FALSE, warning = FALSE, message = FALSE}

library("animation")
library("magrittr")
library("tidyverse")
#library("patchwork") #Nice package for displaying multiple plots together
```


# Central Limit Theorem Simulations

##1. Normally Distributed Random Variables 

First, let's generate data from a normal distribution:

$$X_i \stackrel{i.i.d.}{\sim} N(\mu, \sigma^2)$$

where 

$$\mu = 10 \ \ {\rm and}\ \ \sigma^2 = 36.$$

```{r, echo = F, message = FALSE, warning=FALSE}
#Define mu and sigma
mu <- 10
sigma2 <- 36

# Generate 100 normal random variables from N(10,36) to observe the underlying distribution
N <- 100

# Random sample data frame
set.seed(4)
norm.df <- tibble(
  sample = rnorm(N, mu, sigma2),
  id     = "Original Normal Distribution"
)


#Visualize
norm.df %>%
  ggplot(aes(x = sample, y = ..density.., fill = id)) +
  geom_histogram(binwidth = 10, colour = "black", alpha = 0.1) +
  geom_density(alpha = 0.4) +
  labs(
    x = sprintf("Xi ~ N(%i, %i)", mu, sigma2),
    y = "Density",
    title = "Histogram and Density Curve for the Original Normal Data") +
  scale_fill_viridis_d("") +
  theme(legend.position = "none")
  

```

* What if we can only observe/sample 10 observations from this population?  

* What if we repeatedly took samples of size 10 from the original population?

* How would the means of each smaller sample be distributed, compare to the whole sample?  


Generate and plot the mean sampling distribution of 100 random samples of size 10 (with replacement) from the original Normal distribution

```{r, echo = F, message = FALSE, warning=FALSE}
# Generate 100 random samples of size 10 (with replacement) from the original distribution
n.sample <- 100
n        <- 10

#Sample indices
sample.indices   <- list()


for (i in 1:n.sample) {
  
  sample.indices[[i]] <- sample(1:N, n, replace = TRUE) 
  
}

#Take the mean of each sample and attach to original data frame

norm.df <- bind_rows(norm.df,
                     tibble(
                       sample = map_dbl(.x = sample.indices,
                                        ~mean(norm.df$sample[.x])),
                       id     = "Sample Means Distribution"
                     ))

#Visualize
norm.df %>%
  ggplot(aes(x = sample, y = ..density.., fill = id)) +
  geom_histogram(binwidth = 10, colour = "black", alpha = 0.2) +
  geom_density(alpha = 0.4, adjust = 1.6) +
  labs(
    x = sprintf("Xi ~ N(%i, %i)", mu, sigma2),
    y = "Density") +
  scale_fill_viridis_d("Sample")
```


##2. Non-normally distributed random variables

## Poisson Distribution

Generate data from a Poisson distribution:

$$X_i \stackrel{i.i.d.}{\sim} Pois(2)$$

```{r, echo = F, message = FALSE, warning=FALSE}

# Generate 100 random variables from Pois(5) 
N <- 100
lambda <- 2

# Random sample data frame
set.seed(2)
pois.df <- tibble(
  sample = rpois(N, 5),
  id     = "Original Poisson Distribution"
)
```

Generate and plot the mean sampling distribution of 100 random samples of size 10 (with replacement) from the original Poisson distribution:
```{r, echo = F, message = FALSE, warning=FALSE}

n.sample <- 100
n        <- 10

#Sample indices
sample.indices   <- list()


for (i in 1:n.sample) {
  
  sample.indices[[i]] <- sample(1:N, n, replace = TRUE) 
  
}

#Take the mean of each sample and attach to original data frame

pois.df <- bind_rows(pois.df,
                     tibble(
                     sample = map_dbl(.x = sample.indices,
                                        ~mean(pois.df$sample[.x])),
                       id     = "Sample Means Distribution"
                     ))

#Visualize
pois.df %>%
  ggplot(aes(x = sample, y = ..density.., fill = id)) +
  geom_histogram(bins=30, colour = "black", alpha = 0.2) +
  geom_density(alpha = 0.4) +
  labs(
    x = sprintf("Xi ~ Pois(2)"),
    y = "Density") +
  scale_fill_viridis_d("Sample")
```

Generate and plot the mean sampling distribution of 100 random samples of size 30 (with replacement) from the original Poisson distribution:

```{r, echo = F, message = FALSE, warning=FALSE}

n.sample <- 100
n        <- 30

#Sample indices
sample.indices   <- list()


for (i in 1:n.sample) {
  
  sample.indices[[i]] <- sample(1:N, n, replace = TRUE) 
  
}

#Take the mean of each sample and attach to original data frame

pois.df <- bind_rows(pois.df,
                     tibble(
                     sample = map_dbl(.x = sample.indices,
                                        ~mean(pois.df$sample[.x])),
                       id     = "Sample Means Distribution"
                     ))

#Visualize
pois.df %>%
  ggplot(aes(x = sample, y = ..density.., fill = id)) +
  geom_histogram(bins=30, colour = "black", alpha = 0.2) +
  geom_density(alpha = 0.4) +
  labs(
    x = sprintf("Xi ~ Pois(2)"),
    y = "Density") +
  scale_fill_viridis_d("Sample")
```




## Binomial Distribution

Generate data from a Binomial distribution:

$$X_i \stackrel{i.i.d.}{\sim} Bin(50,0.05)$$

```{r, echo = F, message = FALSE, warning=FALSE}

# Generate 100 random variables from Bin(50,0.30) distribution 
N <- 100
nbin <- 50
p <- 0.05

# Random sample data frame
set.seed(2)
bin.df <- tibble(
  sample = rbinom(N, nbin, p),
  id     = "Original Binomial Distribution"
)
```

Generate and plot the mean sampling distribution of 100 random samples of size 10 (with replacement) from the original Binomial distribution:
```{r, echo = F, message = FALSE, warning=FALSE}

n.sample <- 100
n        <- 10

#Sample indices
sample.indices   <- list()


for (i in 1:n.sample) {
  
  sample.indices[[i]] <- sample(1:N, n, replace = TRUE) 
  
}

#Take the mean of each sample and attach to original data frame

bin.df <- bind_rows(bin.df,
                     tibble(
                     sample = map_dbl(.x = sample.indices,
                                        ~mean(bin.df$sample[.x])),
                       id     = "Sample Means Distribution"
                     ))

#Visualize
bin.df %>%
  ggplot(aes(x = sample, y = ..density.., fill = id)) +
  geom_histogram(bins=30, colour = "black", alpha = 0.2) +
  geom_density(alpha = 0.4) +
  labs(
    x = sprintf("Xi ~ Bin(50,0.05)", mu, sigma2),
    y = "Density") +
  scale_fill_viridis_d("Sample")
```


Generate and plot the mean sampling distribution of 100 random samples of size 30 (with replacement) from the original Binomial distribution:
```{r, echo = F, message = FALSE, warning=FALSE}

n.sample <- 100
n        <- 30

#Sample indices
sample.indices   <- list()


for (i in 1:n.sample) {
  
  sample.indices[[i]] <- sample(1:N, n, replace = TRUE) 
  
}

#Take the mean of each sample and attach to original data frame

bin.df <- bind_rows(bin.df,
                     tibble(
                     sample = map_dbl(.x = sample.indices,
                                        ~mean(bin.df$sample[.x])),
                       id     = "Sample Means Distribution"
                     ))

#Visualize
bin.df %>%
  ggplot(aes(x = sample, y = ..density.., fill = id)) +
  geom_histogram(bins=30, colour = "black", alpha = 0.2) +
  geom_density(alpha = 0.4) +
  labs(
    x = sprintf("Xi ~ Bin(50,0.05)", mu, sigma2),
    y = "Density") +
  scale_fill_viridis_d("Sample")
```











